---
title: "phosphoRWHN_MAPKERK"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{phosphoRWHN_MAPKERK}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(phosphoRWHN)
library(tidyverse)
```
#PhosphoRWHN
Phosphoproteomics data ....
PhosphoRWHN addresses the problem, by providing a systems level analysis method for assigning potential functions to phosphorylated sites. Our manuscript on this work is described [here](www.biorxiv.com)


## Data Preparation

Any phosphorpteomics dataset can be used. The user may wish to filter the data first to remove "unregulated" or "insignificant" sites. However, as a systems-level analysis, the algorithm performs best on data describing different phosphorylation profile, so use all of your filtered data even if you intend to only look at individual sites or clusters in the end

From the phosphoproteomics dataset, we need at least three variables:
* HGNC symbol
* Modified Amino Acid (S, T, Y)
* Amino acid position

These need to be concatenated into the form HGNC_AminoAcid+Position (e.g. MAPK1_T185) to create an identifier that is parsed by the algorithm.


We can use the quantitative data to cluster the sites prior to running RWHN

```{r}
model <- read.csv(
  system.file("extdata", "model_phospho.csv", package = "phosphoRWHN", mustWork = T)
) 
model

model <- model %>% 
  mutate(id = paste0(prot, "_", site)) %>% 
  select(id, t1:t5)

model
```

In this case we are going to use the `mfuzz()` function from the *Mfuzz* package to cluster the data. The data can be clustered in any way. A discussion on clustering methods for phosphoproteomicscan be found [here](www.google.com)
```{r}
library(Mfuzz)

set.seed(1)
inputdata <- model %>% 
    tibble::remove_rownames() %>%
    tibble::column_to_rownames(var = "id") 
phospep <- new("ExpressionSet", 
               exprs = as.matrix(inputdata))
phospep.z <- standardise(phospep)      #Avg expression of each peptide = 0, sd = 1
optimalM <- mestimate(phospep.z)   #set fuzzifier
cl <- mfuzz(phospep.z, c = 5, m = optimalM)

cl$cluster
```

The variable cl$cluster is the input for phosphoRWHN.

## Running the pipeline

The package has three functions that, run sequentially, form the analysis pipeline.
1. Construct the heterogeneous network - `constructHetNet()`
2. Run RWHN - `calculateRWHN()`
3. Filter & visualise - `heatmap_RWHN()`

### Construct the heterogeneous network
The function `constructHetNet` takes several parameters which are detailed in the documentation. Parameters without default parameters are:
* `clustering` - in our case, cl$cluster
* `phosphoData` - this is optional, and is used to trim edges in the phospho-layer of the network. In this case we use our processed *model* data frame.

The output is a list containing the edgelists and vertices of the multilayer network.

```{r}

# 1. Construct the heterogeneous network
hetNet <- constructHetNet(clustering = cl$cluster, stringConf = 0.4, modules = T)

# 2. Run RWHN, using sites in each cluster as seeds
seed_l <- lapply(1:max(cl$cluster), function(i){
  c(names(cl$cluster[cl$cluster == i]))
})
set.seed(1)
rwhn <- lapply(seed_l, function(s){
  calculateRWHN(hetNet = hetNet,
                seeds = s, 
                transitionProb = 0.7, 
                restart = 0.7,
                eta_xy =  0.3,
                eta_yz = 0.7,
                 eps = 1/10^12
                )
})

# 3. Filter output & visualise
sighm <-
  heatmap_RWHN(rwhn,
               database = "GOBP",
               colours = c(low = "#ffe6e8", high = "#f74451"))

sighm

ora <- overrepresentationAnalysis(clustering = cl$cluster, RWHN_sig = sighm, vis = T)
ora
```

## Statistical Assessment

To reproduce the Figure S2B in the paper, randomly permuted networks need to be generated.
This can take some time depending on the size of your network.

```{r eval = F}
rwhn_random <- lapply(seed_l, function(s) {
  calculateRWHN(
    hetNet = hetNet,
    seeds = s,
    transitionProb = 0.7,
    restart = 0.7,
    eta_xy =  0.3,
    eta_yz = 0.7,
    eps = 1 / 10 ^ 12,
    random = T
  )
})

# Generate results lists of non-random data with different cut-offs
sighm_1 <- heatmap_RWHN(
  rwhn,
  #, ylab = "GOBP Term",
  colours = c(low = "#ffe6e8", high = "#f74451"),
  removeCommon = T,
  pct_cutoff = 0.01
) %>%
  .$data %>%
  mutate(pct_cutoff = 0.01)
sighm_5 <- heatmap_RWHN(
  rwhn,
  colours = c(low = "#ffe6e8", high = "#f74451"),
  removeCommon = T,
  pct_cutoff = 0.05
) %>%
  .$data %>%
  mutate(pct_cutoff = 0.05)
sighm_10 <- heatmap_RWHN(
  rwhn,
  #, ylab = "GOBP Term",
  colours = c(low = "#ffe6e8", high = "#f74451"),
  removeCommon = T,
  pct_cutoff = 0.1
) %>%
  .$data %>%
  mutate(pct_cutoff = 0.1)
sighm_15 <- heatmap_RWHN(
  rwhn,
  #, ylab = "GOBP Term",
  colours = c(low = "#ffe6e8", high = "#f74451"),
  removeCommon = T,
  pct_cutoff = 0.15
) %>%
  .$data %>%
  mutate(pct_cutoff = 0.15)

# Calculate p-values
ptests <-
  lapply(1:length(seed_l), function(i) {
    lapply(v$v[v$layer == "func"], function(term) {
      ranks <-  sapply(rwhn_random[[i]],
                       function(x)
                         ifelse(term %in% x$name, x$rank[x$name == term], NA))
      
      if (all(is.na(ranks))) {
        return(data.frame())
      }
      
      true_rank <- rwhn[[i]]$rank[rwhn[[i]]$name == term]
      t_p <- t.test(ranks, mu = true_rank)$p.value
      t_padj <- p.adjust(t_p, "fdr")
      mann_p <- wilcox.test(ranks, mu = true_rank)$p.value
      mann_padj <- p.adjust(mann_p, "fdr")
      
      res <- data.frame(term,
                        true_rank,
                        true_prob =  rwhn[[i]]$V1[rwhn[[i]]$name == term],
                        t_p,
                        t_padj,
                        mann_p,
                        mann_padj)
      return(res)
    }) %>%
      bind_rows() %>%
      mutate(
        sig_t = ifelse(t_padj < 0.05, T, F),
        sig_mann = ifelse(mann_padj < 0.05, T, F),
        insighm_1 = ifelse(term %in% sighm_1[sighm_1$seed == i, ]$name, T, F),
        insighm_5 = ifelse(term %in% sighm_5[sighm_5$seed == i, ]$name, T, F),
        insighm_10 = ifelse(term %in% sighm_10[sighm_10$seed == i, ]$name, T, F),
        insighm_15 = ifelse(term %in% sighm_15[sighm_15$seed == i, ]$name, T, F),
        cluster = i
      ) %>%
      arrange(true_rank)
  }) %>%
  bind_rows()

# Visualise
mannwhitney_hm <- ptests %>%
  pivot_longer(
    c(sig_mann, insighm_1, insighm_5,
      insighm_10, insighm_15),
    names_to = "test",
    values_to = "p"
  ) %>%
  ggplot(aes(
    x = factor(
      test,
      levels = c(
        "sig_mann",
        "insighm_1",
        "insighm_5",
        "insighm_10",
        "insighm_15"
      )
    ),
    y = term,
    fill = p
  )) +
  geom_tile() +
  scale_x_discrete(
    name = "",
    labels = c("Mann Whitney-U\n< 0.05",
               "1%",
               "5%",
               "10%",
               "15%")
  ) +
  xlab("GOBP Term") +
  scale_fill_manual(name = "",
                    values = c("TRUE" = "#1ecbe1", "FALSE" = "#E1341E")) +
  facet_wrap( ~ cluster, nrow = 1) +
  theme(
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.title = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.text.x = element_text(
      size = 5,
      angle = 45,
      hjust = 1
    ),
    legend.key.height = unit(0.25, "cm"),
    legend.key.width = unit(0.25, "cm"),
    legend.text = element_text(size = 5),
    plot.margin = margin(0, 0, 0, 0, "cm"),
    legend.margin = margin(0, 0, 0, 0, "cm")
  )


mannwhitney_hm

```



